---
title: "Some of the Answers to Some of Your Network Questions"
linktitle: "Week 1: Some of the Answers to Some of Your Network Questions"
menu:
  readings:
    parent: Readings
    weight: 1
type: docs
weight: 1
bibliography: "../../static/bib/references.bib"
csl: "../../static/bib/chicago-fullnote-bibliography-no-bib.csl"
output:
  blogdown::html_page:
    toc: true
---

```{r load_packages, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(fontawesome)
```

### Get Some Advice
First things first:
<br>
<br>
<iframe src="/videos/snoop_syllabus.mp4" width="320" height="480" frameBorder="0">Read the syllabus.</iframe>
<br>
<br>Far be it from us to disagree with Snoop, so you should probably go read the [syllabus](/syllabus/) as soon as possible!

### Review Syllabus
Review the overviews of the other areas as well: [readings](/readings/), [data camp](/datacamp/), [materials](/materials/), [tasks](/tasks/), [to do list](/due/), and [resources](/resources/) pages for this class.

### Book Stuff
```{r icontibble, echo=FALSE, message=FALSE}
book <- tibble(
  
  chapter <- c("1"),
  
  title <- c("The What and the Why of Statistics"),
  
  text <- c("Social Statistics for a Diverse Society"),
  
  pages <- c("1-19"),
  
  .name_repair = "minimal"
  
)
```

<center>
```{r, echo=FALSE}
book %>%
  kbl(col.names = c("Chapter", "Title", "Source", "Pages"), 
      "html", 
      escape = FALSE,
      align = 'llll') %>%
  column_spec(1, 
              bold = TRUE)
```
</center>
