---
title: "Measures of Central Tendency"
linktitle: "Week 3: Measures of Central Tendency Content"
output:
  blogdown::html_page:
    toc: true
menu:
  materials:
    parent: Content
    weight: 3
type: docs
weight: 2
editor_options: 
  chunk_output_type: console
---

### Textbook Notes

- [Chapter 3 PowerPoint](/lecture_notes/Week 3/SSDS_Ch3.pptx){target="_blank"} 
- [Chapter 3 Lecture Notes](/lecture_notes/Week 3/SSDS_Ch3.docx){target="_blank"}

### In Class Notes

Click on the presentation itself and then you may

- Use this [link](/slides/Week 3/Slides-Week-3.html){target="_blank"} to view a larger version of slideshow in a new window.

```{r echo = FALSE}
knitr::include_url("/slides/Week 3/Slides-Week-3.html")
```

You can move back and forth using <kbd>←</kbd> and <kbd>→</kbd> and toggle fullscreen by pressing <kbd>F</kbd>. Just press <kbd>O</kbd>^[For **O**verview] at any point in your slideshow and a display of all slides, called tile view, will appear. Click on a slide to go there or press <kbd>O</kbd> to exit tile view. For more options, press <kbd>H</kbd> for help.

You can download 

- a [PDF](/slides/Week 3/Slides-Week-3.pdf){target="_blank"} of the presentation. 

<!--
### R Walkthrough

Click on the presentation itself and then you may

- Move back and forth through the slideshow using the ⬅ and ⮕ buttons on your keyboard.
- Press the letter **O** at any point to see see a tile view of the slideshow.
- Use this [link](/slides/Week 2/Slides-Week-2R.html){target="_blank"} to view a larger version of slideshow in a new window.

```{r echo = FALSE}
knitr::include_url("/slides/Week 2/Slides-Week-2R.html")
```

You can download 

- a [PDF](/slides/Week 2/Slides-Week-2R.pdf){target="_blank"} of the presentation above. 
- a [script](/slides/Week 2/Slides-Week-2R.R){target="_blank"} file of just the R chunks used in the presentation.
 -->