---
title: "The Not so Tiny t-test"
linktitle: "Week 10: The Not so Tiny t-test"
author: "Week 10"
output: 
 xaringan::moon_reader:
   css: xaringan-themer.css
   nature:
     ratio: 16:9
     highlightStyle: github
     highlightLines: true
     countIncrementalSlides: false
     navigation:
         scroll: false
---

<script src="https://ajax.googleapis.com/ajax/libs/jquery/3.6.0/jquery.min.js"></script>

<script type="text/x-mathjax-config">
MathJax.Hub.Register.StartupHook("TeX Jax Ready",function () {
  MathJax.Hub.Insert(MathJax.InputJax.TeX.Definitions.macros,{
    cancel: ["Extension","cancel"],
    bcancel: ["Extension","cancel"],
    xcancel: ["Extension","cancel"],
    cancelto: ["Extension","cancel"]
  });
});
</script>

<style>
section {
    display: flex;
    display: -webkit-flex;
}

section p {
    margin: auto;
}

.hljs-github .hljs {
    background: transparent;
    color: #b2dfdb;
}

.hljs-github .hljs-keyword {
    color: #64b5f6;
}

.hljs-github .hljs-literal {
    color: #64b5f6;
}

.hljs-github .hljs-number {
    color: #64b5f6;
}

.hljs-github .hljs-string {
    color: #b7b3ef;
}

section {
    height: 600px;
    width: 60%;
    margin: auto;
    border-radius: 20px;
    background-color: #212121;
}

section p {
    text-align: center;
    font-size: 30px;
    background-color: #212121;
    border-radius: 20px;
    font-family: Roboto Condensed;
    font-style: bold;
    padding: 15px;
    color: #bff4ee;
}

#center {
text-align: center;
}

.center p {
  margin: 0;
  position: absolute;
  top: 50%;
  left: 50%;
  -ms-transform: translate(-50%, -50%);
  transform: translate(-50%, -50%);

</style>

```{r setup, include=FALSE, purl=FALSE}
library(tidyverse)
library(fontawesome)
library(knitr)
library(kableExtra)
library(patchwork)
library(flipbookr)
```

```{r echo = FALSE, purl=FALSE}
xaringanthemer::style_duo(
  primary_color = "#212121",
  secondary_color = "#f4eebf",
  code_inline_background_color = "transparent",
  code_inline_color = "#b2dfdb",
  table_row_border_color = "#212121",
  table_row_even_background_color = "#212121",
  footnote_font_size = "0.6em",
  header_font_google = xaringanthemer::google_font("Roboto Condensed", "700"),
  text_font_google   = xaringanthemer::google_font("Roboto Condensed", "400")
)

xaringanExtra::use_xaringan_extra(c("tile_view", "animate_css", "tachyons"))

xaringanExtra::use_logo(
  image_url = here::here("static", "img", "course_hex_alpha.png"),
  link_url = "https://edp613.asocialdatascientist.com",
  position = xaringanExtra::css_position(top = "1em", right = "1em")
)

opts_chunk$set(dev.args=list(bg="transparent"))
```

# Packages needed and a Note about Icons

Please load up the following packages. Remember to first install the ones you don't have.
```{r echo = FALSE, eval = TRUE, message=FALSE}
library(tidyverse)
library(mosaic)
library(ggplot2movies)
``` 

```{r echo = FALSE, eval = TRUE, message=FALSE}
suppressMessages(library(tidyverse))
suppressMessages(library(mosaic))
suppressMessages(library(ggplot2movies))
``` 
<br>
<br>
You may come across the following icons. The table below lists what each means.
```{r eval = TRUE, echo = FALSE, purl=FALSE}
forward <- as.character(fontawesome::fa("forward", fill = "
#4682b4"))

stop <- as.character(fontawesome::fa("stop", fill = "#ff6347"))

link <- as.character(fontawesome::fa("link", fill = "#5cb85c"))

bookmark <- as.character(fontawesome::fa("bookmark", fill = "#faffbd"))


icon_desc <- tibble(
  
  Icon = c(forward,
           stop, 
           link,
           bookmark),
  
  Description = c("Indicates that an example continues on the following slide.",
                  "Indicates that a section using common syntax has ended.",
                  "Indicates that there is an active hyperlink on the slide.",
                  "Indicates that a section covering a concept has ended.")
  
)
```

```{r message=FALSE, warning=FALSE, eval = TRUE, echo = FALSE, purl=FALSE}
kable(icon_desc, 
      escape = FALSE,
      align = 'cl') %>%
  kable_styling(full_width = FALSE) %>%
  column_spec(1, width = "10em") %>%
  column_spec(2, width = "40em") %>%
  row_spec(0, background = "transparent") %>%
  row_spec(1, background = "#transparent") %>%
  row_spec(2, background = "#transparent") %>%
  row_spec(3, background = "#transparent") %>%
  row_spec(4, background = "#transparent") 
```

---

# Comparing the Means Between Groups of Things

The $t$-test is:

--

- One of the most common tests in statistics

--

- Used to determine whether the means of two groups are equal

---

# Ideas

--

> **One-sample *t*-tests**: Compare the sample mean with a known value, when the variance of the population is unknown

--

> **Two-sample *t*-tests**: Compare the means of two groups under the assumption that both samples are random, independent, and normally distributed with unknown but equal variances

--

> **Paired *t*-tests**: Compare the means of two sets of paired samples, taken from two populations with unknown variance

---

## Packages

Please load up the following packages

```{r echo = TRUE, eval = TRUE}
library(tidyverse)
library(patchwork)
```

---

## The `t.test` Command

```{r ttest_definition, eval=FALSE}
t.test(x, y = NULL, 
       alternative = c("two.sided", "less", "greater"), 
       mu = 0, 
       paired = FALSE, 
       var.equal = FALSE, 
       conf.level = 0.95)
```

```{r echo=FALSE}

tibble(

option = c("x", "y", "mu", "alternative", "paired", "var.equal", "conf.level"),

what <- c("a numeric vector from a data set",
          "an optional numeric vector from a data set",
          "a number indicating the true value of the mean",
          "preference on type of test you wish to run",
          "preference on whether you wish to perform a paired <i>t</i>-test",
          "indicates whether or not to assume equal variances when performing a two-sample  <i>t</i>-test",
          "the confidence level of the reported confidence interval")

) %>%
  kbl(col.names = c("Option", "Function"), 
      "html", 
      escape = FALSE,
      align = 'cl') %>%
  kable_styling(full_width = FALSE) %>%
  column_spec(1, width = "10em", 
              color = "#b2dfdb", 
              monospace = TRUE) %>%
  column_spec(2, width = "40em")

```

---

### Note

If `y` is

– excluded, `t.test` will run as a one-sample *t*-test

– included, `t.test` will run as a two-sample *t*-test

- The default `t.test` command will run as a two-sided *t*-test BUT you can perform an alternative
hypothesis by changing the `alternative` option using `greater` or `less`

`t.test(x, alternative = "greater", mu = 47)`

---
